---
title: "OysterCatalyst-SIA"
author: "XXX"
date: "2026-07-21"
output: html_document
---

This notebook compiles the stable isotope mixing-model workflow used for journal submission, including isotope biplots, MixSIAR runs by site, contribution summaries, environmental associations, hypervolume analyses, and trophic-level comparisons.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  error = FALSE,
  fig.width = 14,
  fig.height = 10,
  out.width = "75%"
)

# Clear the workspace and graphics devices before running the analysis.
rm(list = ls())
graphics.off()
# Load the majority of relevant packages
library(tidyverse) # For data manipulation and summarization
library(plotrix) #To use standard error calculation
library(cowplot) #For grouping plots
library(scales) #For editing axis scales
library(vegan) #For multivariate analyses
library(rstatix) #For running ANCOVA
library(ggpmisc) #adding equation and r2 to figures

alpha <- 0.05
formula <- y ~ x  # Generic y~x formula placeholder used in plotting helpers.
theme_set(theme_bw(base_size = 20))
```

## Stable isotope biplots

Create carbon-nitrogen and carbon-sulfur biplots for source and consumer isotope means with ±1 SD error bars across sites.

```{r}
#Read in source data for all sites
#Rename site and source names for consistency
sia<-read.csv("Data/SIA/MCsources.csv")%>%mutate(Site = fct_recode(Site, "NI-WB"="NIWB"))%>%mutate(Source = fct_recode(Source, "Benthic microalgae"="Benthic algae", "S. alterniflora"="Spartina", "sPOM" = "POM"))

#Order sites from north to south
sia$Site<-factor(sia$Site, levels=c("NC", "NI-WB", "SI", "GTM"))

#Create C&N biplot
cn<-ggplot(sia, aes(x=Meand13C, y=Meand15N))+
  geom_errorbar(aes(ymax=Meand15N+SDd15N, ymin=Meand15N-SDd15N))+
                  geom_errorbarh(aes(xmax=Meand13C+SDd13C, xmin=Meand13C-SDd13C))+
  geom_point(aes(shape=Source,fill=Site),color="black", size=4)+
  scale_shape_manual(values=c(21,22,23,24,25))+
   scale_fill_manual(values=c("#e66101", "#fdb863","#B2abd2","#5e3c99"))+
  guides(fill=guide_legend(override.aes=list(shape=21)))+
  theme(plot.title = element_text(size = 16, face = "bold"))+
  labs(x=expression(delta*"13C"), y=expression(delta*"15N"))+
    theme(panel.border = element_rect(colour = "black", fill=NA, size=1))+theme(legend.position="none")

#Create C & S biplot
cs<-ggplot(sia, aes(x=Meand13C, y=Meand34S))+
  geom_errorbar(aes(ymax=Meand34S+SDd34S, ymin=Meand34S-SDd34S))+
                  geom_errorbarh(aes(xmax=Meand13C+SDd13C, xmin=Meand13C-SDd13C))+
  geom_point(aes(shape=Source,fill=Site),color="black", size=4)+
  scale_shape_manual(values=c(21,22,23,24,25))+
  scale_fill_manual(values=c("#e66101", "#fdb863","#B2abd2","#5e3c99"))+
  guides(fill=guide_legend(override.aes=list(shape=21)))+
  theme(plot.title = element_text(size = 16, face = "bold"))+
  labs(x=expression(delta*"13C"), y=expression(delta*"34S"))+
    theme(panel.border = element_rect(colour = "black", fill=NA, size=1))

#Use package cowplot to combine figures 
library(cowplot)
plot_grid(cn, cs, rel_widths = c(0.7, 1))

#Read in consumer data for all sites 
cons<-read.csv("Data/SIA/AllConsumers.csv")%>%
  group_by(Site, Species)%>%
  summarize(mdC=mean(d13C), sdC=sd(d13C),mdN=mean(d15N), sdN=sd(d15N), mdS=mean(d34S), sdS=sd(d34S), tot=n())%>%
  mutate(Site = fct_recode(Site, "NI-WB"="NIWB"))

#Reorder sites from north to south
cons$Site<-factor(cons$Site, levels=c("NC", "NI-WB", "SI", "GTM"))

#create C & N biplot
cn1<-ggplot(cons, aes(x=mdC, y=mdN))+
  geom_errorbar(aes(ymax=mdN+sdN, ymin=mdN-sdN))+
                  geom_errorbarh(aes(xmax=mdC+sdC, xmin=mdC-sdC))+
  geom_point(aes(shape=Species,fill=Site),color="black", size=4)+
  scale_shape_manual(values=c(21,22,23,24,25))+
   scale_fill_manual(values=c("#e66101", "#fdb863","#B2abd2","#5e3c99"))+
  guides(fill=guide_legend(override.aes=list(shape=21)))+
  theme(plot.title = element_text(size = 16, face = "bold"))+
  labs(x=expression(delta*"13C"), y=expression(delta*"15N"))+
    theme(panel.border = element_rect(colour = "black", fill=NA, size=1))+theme(legend.position="none")

#Create C & S Biplot
cs1<-ggplot(cons, aes(x=mdC, y=mdS))+
  geom_errorbar(aes(ymax=mdS+sdS, ymin=mdS-sdS))+
                  geom_errorbarh(aes(xmax=mdC+sdC, xmin=mdC-sdC))+
  geom_point(aes(shape=Species,fill=Site),color="black", size=4)+
  scale_shape_manual(values=c(21,22,23,24,25))+
  scale_fill_manual(values=c("#e66101", "#fdb863","#B2abd2","#5e3c99"))+
  guides(fill=guide_legend(override.aes=list(shape=21)))+
  theme(plot.title = element_text(size = 16, face = "bold"))+
  labs(x=expression(delta*"13C"), y=expression(delta*"34S"))+
    theme(panel.border = element_rect(colour = "black", fill=NA, size=1))

#Combine figures
plot_grid(cn1, cs1, rel_widths = c(0.7, 1), ncol=2)

#Save file of consumer mean (used to create tables)
#write.csv(cons, "Data/SIA/AllConsumerMeans.csv")
```

## MixSIAR mixing models

```{r}
#Open MixSIAR package used for stable isotope mixing model
library(MixSIAR)
```

### Run mixing model for Sapelo Island data (Georgia)
```{r}
#Read in consumer data
mix1 <- load_mix_data(filename="Data/SIA/MCconsumersSI.csv",
                      iso_names=c("d13C", "d34S"),
                      factors=c("Species"),
                      fac_random=c(FALSE),
                      fac_nested=c(FALSE),
                      cont_effects=NULL)

#Read in source data
source1 <- load_source_data(filename="Data/SIA/MCsourcesSI.csv",
                            source_factors = NULL,
                            conc_dep=FALSE, 
                            data_type="means", 
                            mix1)

#Read in trophic discrimination factors
discr1 <- load_discr_data(filename="Data/SIA/MCdiscrimination.csv", mix1)

#Plot data into isotopic space
plot_data(filename="Outputs/SIA/SI/isospace_plot",
          plot_save_pdf=FALSE,
          plot_save_png=TRUE,
          mix1,source1,discr1)


# Define model structure and write JAGS model file
model_filename <- "Outputs/SIA/SI/MixSIAR_model_SI.txt"
resid_err <- FALSE
process_err <- TRUE
write_JAGS_model(model_filename, resid_err, process_err, mix1, source1)

# Run the JAGS model
#jagsNC.1 <- run_model(run="test", mix, source, discr, model_filename, alpha.prior=1, resid_err, process_err)
jagsSI <- run_model(run="normal", mix1, source1, discr1, model_filename, alpha.prior=1, resid_err, process_err)

# Process diagnostics, summary stats, and posterior plots
output_options = list(summary_save = TRUE,
                      summary_name = "Outputs/SIA/SI/SI_summary_statistics1",
                      sup_post = T,
                      plot_post_save_pdf = F,
                      plot_post_name = "Outputs/SIA/SI/SI_posterior_density1",
                      sup_pairs = T,
                      plot_pairs_save_pdf = F,
                      plot_pairs_name = "Outputs/SIA/SI/SI_pairs_plot1",
                      sup_xy = T,
                      plot_xy_save_pdf = FALSE,
                      plot_xy_name = "Outputs/SIA/SI/SI_xy_plot1",
                      gelman = TRUE,
                      heidel = FALSE,
                      geweke = TRUE,
                      diag_save = T,
                      diag_name = "Outputs/SIA/SI/SI_diagnostics",
                      indiv_effect = FALSE,
                      plot_post_save_png = FALSE,
                      plot_pairs_save_png = FALSE,
                      plot_xy_save_png = FALSE,
                      diag_save_ggmcmc = FALSE)

# display outputs for SI model 
output_JAGS(jagsSI, mix1, source1, output_options)


```


### Run mixing model for GTM data (Florida)

```{r}
#Read in consumer data
mix2 <- load_mix_data(filename="Data/SIA/MCconsumersGTM.csv",
                      iso_names=c("d13C", "d34S"),
                      factors=c("Species"),
                      fac_random=c(FALSE),
                      fac_nested=c(FALSE),
                      cont_effects=NULL)
#Read in source data
source2 <- load_source_data(filename="Data/SIA/MCsourcesGTM.csv",
                            source_factors = NULL,
                            conc_dep=FALSE, 
                            data_type="means", 
                            mix2)

#Read in discrimination factors
discr2 <- load_discr_data(filename="Data/SIA/MCdiscrimination.csv", mix2)

#Plot data into isotopic space
plot_data(filename="Outputs/SIA/GTM/isospace_plot",
          plot_save_pdf=FALSE,
          plot_save_png=TRUE,
          mix2,source2,discr2)


# Define model structure and write JAGS model file
model_filename1 <- "Outputs/SIA/GTM/MixSIAR_model_GTM.txt"
resid_err <- FALSE
process_err <- TRUE
write_JAGS_model(model_filename1, resid_err, process_err, mix2, source2)

# Run the JAGS model ("test" first, then "short")
#jagsNC.1 <- run_model(run="test", mix, source, discr, model_filename, alpha.prior=1, resid_err, process_err)
jagsGTM <- run_model(run="normal", mix2, source2, discr2, model_filename1, alpha.prior=1, resid_err, process_err)

# Process diagnostics, summary stats, and posterior plots
#output_JAGS(jagsNC.1, mix, source)

output_options2 = list(summary_save = TRUE,
                      summary_name = "Outputs/SIA/GTM/GTM_summary_statistics1",
                      sup_post = T,
                      plot_post_save_pdf = F,
                      plot_post_name = "Outputs/SIA/GTM/GTM_posterior_density1",
                      sup_pairs = T,
                      plot_pairs_save_pdf = F,
                      plot_pairs_name = "Outputs/SIA/GTM/GTM_pairs_plot1",
                      sup_xy = TRUE,
                      plot_xy_save_pdf = FALSE,
                      plot_xy_name = "Outputs/SIA/GTM/GTM_xy_plot1",
                      gelman = TRUE,
                      heidel = FALSE,
                      geweke = TRUE,
                      diag_save = T,
                      diag_name = "Outputs/SIA/GTM/GTM_diagnostics",
                      indiv_effect = FALSE,
                      plot_post_save_png = FALSE,
                      plot_pairs_save_png = FALSE,
                      plot_xy_save_png = FALSE,
                      diag_save_ggmcmc = FALSE)
# display outputs
output_JAGS(jagsGTM, mix2, source2, output_options2)


```


### Run mixing model for NI-WB data (South Carolina)

```{r}
#Read in consumer data
mix3 <- load_mix_data(filename="Data/SIA/MCconsumersNI.csv",
                      iso_names=c("d13C", "d34S"),
                      factors=c("Species"),
                      fac_random=c(FALSE),
                      fac_nested=c(FALSE),
                      cont_effects=NULL)
#Read in source data
source3 <- load_source_data(filename="Data/SIA/MCsourcesNI.csv",
                            source_factors = NULL,
                            conc_dep=FALSE, 
                            data_type="means", 
                            mix3)
#Read in trophic discrimination factors
discr3 <- load_discr_data(filename="Data/SIA/MCdiscrimination.csv", mix3)

#Plot data in isotopic space
plot_data(filename="Outputs/SIA/NI/isospace_plot",
          plot_save_pdf=FALSE,
          plot_save_png=TRUE,
          mix3,source3,discr3)


# Define model structure and write JAGS model file
model_filename3 <- "Outputs/SIA/NI/MixSIAR_model_NI.txt"
resid_err <- FALSE
process_err <- TRUE
write_JAGS_model(model_filename3, resid_err, process_err, mix3, source3)

# Run the JAGS model ("test" first, then "short")
#jagsNC.1 <- run_model(run="test", mix, source, discr, model_filename, alpha.prior=1, resid_err, process_err)
jagsNI <- run_model(run="normal", mix3, source3, discr3, model_filename3, alpha.prior=1, resid_err, process_err)

# Process diagnostics, summary stats, and posterior plots
#output_JAGS(jagsNC.1, mix, source)

output_options3 = list(summary_save = TRUE,
                      summary_name = "Outputs/SIA/NI/NI_summary_statistics1",
                      sup_post = T,
                      plot_post_save_pdf = F,
                      plot_post_name = "Outputs/SIA/NI/NI_posterior_density1",
                      sup_pairs = T,
                      plot_pairs_save_pdf = F,
                      plot_pairs_name = "Outputs/SIA/NI/NI_pairs_plot1",
                      sup_xy = TRUE,
                      plot_xy_save_pdf = FALSE,
                      plot_xy_name = "Outputs/SIA/NI/NI_xy_plot1",
                      gelman = TRUE,
                      heidel = FALSE,
                      geweke = TRUE,
                      diag_save = T,
                      diag_name = "Outputs/SIA/NI/NI_diagnostics",
                      indiv_effect = FALSE,
                      plot_post_save_png = FALSE,
                      plot_pairs_save_png = FALSE,
                      plot_xy_save_png = FALSE,
                      diag_save_ggmcmc = FALSE)
# display outputs for NI-WB
output_JAGS(jagsNI, mix3, source3, output_options3)

```

### Run mixing model for NC data (North Carolina)

```{r}
#Read in consumer data
mix4 <- load_mix_data(filename="Data/SIA/MCconsumersNC.csv",
                      iso_names=c("d13C", "d34S"),
                      factors=c("Species"),
                      fac_random=c(FALSE),
                      fac_nested=c(FALSE),
                      cont_effects=NULL)
#Read in source data
source4 <- load_source_data(filename="Data/SIA/MCsourcesNC.csv",
                            source_factors = NULL,
                            conc_dep=FALSE, 
                            data_type="means", 
                            mix4)
#RTead in trophic discrimination factors
discr4 <- load_discr_data(filename="Data/SIA/MCdiscrimination.csv", mix4)

#Plot data in isotopic space
plot_data(filename="Outputs/SIA/NC/isospace_plot",
          plot_save_pdf=FALSE,
          plot_save_png=TRUE,
          mix4,source4,discr4)


# Define model structure and write JAGS model file
model_filename4 <- "Outputs/SIA/NC/MixSIAR_model_NC.txt"
resid_err <- FALSE
process_err <- TRUE
write_JAGS_model(model_filename4, resid_err, process_err, mix4, source4)

# Run the JAGS model ("test" first, then "short")
#jagsNC.1 <- run_model(run="test", mix, source, discr, model_filename, alpha.prior=1, resid_err, process_err)
jagsNC <- run_model(run="normal", mix4, source4, discr4, model_filename4, alpha.prior=1, resid_err, process_err)

# Process diagnostics, summary stats, and posterior plots
#output_JAGS(jagsNC.1, mix, source)

output_options4 = list(summary_save = TRUE,
                      summary_name = "Outputs/SIA/NC/NC_summary_statistics1",
                      sup_post = T,
                      plot_post_save_pdf = F,
                      plot_post_name = "Outputs/SIA/NC/NC_posterior_density1",
                      sup_pairs = T,
                      plot_pairs_save_pdf = F,
                      plot_pairs_name = "Outputs/SIA/NC/NC_pairs_plot1",
                      sup_xy = TRUE,
                      plot_xy_save_pdf = FALSE,
                      plot_xy_name = "Outputs/SIA/NC/NC_xy_plot1",
                      gelman = TRUE,
                      heidel = FALSE,
                      geweke = TRUE,
                      diag_save = T,
                      diag_name = "Outputs/SIA/NC/NC_diagnostics",
                      indiv_effect = FALSE,
                      plot_post_save_png = FALSE,
                      plot_pairs_save_png = FALSE,
                      plot_xy_save_png = FALSE,
                      diag_save_ggmcmc = FALSE)

# display outputs for NC model
output_JAGS(jagsNC, mix4, source4, output_options4)

```

## Bootstrapped contribution boxplots

Construct bootstrap-based boxplots from model means, standard deviations, and 95% intervals by sampling 100 random individuals within each species, flow regime, and season.

```{r}
library(truncnorm)
#Read in a combined output (summary statistics) from all 4 mixing models above
comb<-read.csv("Data/SIA/NERRCat_summary_statisticsV4.csv")

#Filter out relevant column for bootstrapping the data to 100 individuals
comb1<-comb%>%
  select(Species, Source, Mean, SD, X2.50., X97.50.)

#eco is name of data frame, n = number of samples you want. 
#Create function for bootstraping data using means and credible intervals for estimate source contributions for 100 individuals of each taxa at each site
HVvalues = function(eco,n){
  set.seed(14)
  df_tot = data.frame(Species = character(0),
                      Source = character(0), Vals = integer(0))
  for (i in 1:nrow(eco)){
    Vals = truncnorm::rtruncnorm(n, eco[i,5], eco[i,6], eco[i,3], eco[i,4])
    #Vals = rnorm(n, eco[i,4], eco[i,5])
    Species = list(rep(eco[i,1],n))
    Source = list(rep(eco[i,2],n))
    num = list(seq(1, n, 1))
    df = as.data.frame(c(Species = Species, num = num,
                         Source = Source, as.data.frame(Vals)))
    df_tot = rbind(df_tot, df)
  }
  df_com = tidyr::spread(df_tot, 'Source', 'Vals')
  ####### part below - divides all source contributions by sum of all source conts so that it = 1
  set<-df_com[3:ncol(df_com)]
  set$sum<-rowSums(set)
  cols<-seq(3,ncol(df_com))
  for (i in 1:nrow(df_com)){
    for(j in 1:length(cols)){
      df_com[i,cols[j]]<-df_com[i,cols[j]]/set[i,ncol(set)] #change value here to 1 + number of sources for your data
    }}
  df_com$sum<-rowSums(df_com[3:ncol(df_com)])
  return(df_com)
}

#Use HV valees function (created above) to bootstrap data 
comb100<-HVvalues(comb1, 100)

#Convert data frame from wide format into long format for analysis 
comb100X<-comb100 %>% separate(Species,into=c('Site', 'Species'), sep="-")%>%
  select(Site, Species, `Benthic algae`, `POM`,  `Spartina`)%>%
  gather(Source, Contr, -Species, -Site)

#Reorder sites and species for plotting
comb100X$Site<-factor(comb100X$Site, levels=c("NC", "NI", "SI", "GTM"))
levels(comb100X$Site)<-c("NC", "NIWB", "SI", "GTM")
comb100X$Species<-factor(comb100X$Species, levels=c("Oyster", "Mud crab", "White shrimp", "Blue crab", "Pinfish"))


#Create a file to examine the ratios between different sources at each site
combRatio<-comb100 %>% separate(Species,into=c('Site', 'Species'), sep="-")%>%
  select(Site, Species, `Benthic algae`, `POM`)

#Create ratio column berween sPOM and benthic microalgae
combRatio$POMBAratio<-(combRatio$POM-combRatio$`Benthic algae`)

#Reorder sites and speices
combRatio$Site<-factor(combRatio$Site, levels=c("NC", "NI", "SI","GTM"))
levels(combRatio$Site)<-c("NC", "NIWB", "SI", "GTM")
combRatio$Species<-factor(combRatio$Species, levels=c("Oyster", "Mud crab", "White shrimp", "Blue crab", "Pinfish"))

#Save bootstrapped data into CSV file for later use
#write.csv(comb100X, "comb100.csv")

#Plot the proportional contributions of basal resources for each taxa and site

#Rename sites and sources for consistency
comb100X<-comb100X%>%mutate(Site = fct_recode(Site, "NI-WB"="NIWB"))%>%mutate(Source = fct_recode(Source, "Benthic microalgae"="Benthic algae", "S. alterniflora"="Spartina", "sPOM" = "POM"))

#Plot the proportional contributions of basal resources for each taxa and site
ggplot(comb100X, aes(x=Source, y=Contr))+
  geom_boxplot(aes(fill=Source))+
  scale_fill_brewer(palette="Purples")+
  facet_grid(Site~Species)+
  theme_bw()+ 
  labs(y="Proportional contribution", x="Basal resource")+
  theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position="none")
  
```

## Examine the relationships between source contributions and environmental characteristics

```{r}
#Read in environmetal data characteristic data
char2<-read.csv("Data/AllNERRCatDataSummarized.csv")%>%
  group_by(RESERVE)%>%
  summarize_all(list(mean))%>%
  rename(Site=RESERVE)

#Select relevant columns 
char1<-char2[,c(4:9, 13,17,21,25,38,39,40)]

#Look at correlation matrix to determine correlated parameters
corrplot::corrplot.mixed(cor(char1))

#Combine bootstrap contribution data to reserve characteristics
combenvir<-left_join(char2, combRatio, by="Site")
combenvir<-combenvir[,-c(2,3)]

combenvir$Site<-factor(combenvir$Site, levels=c("NC", "NIWB", "SI", "GTM"))
```


```{r}
#Test correlation between PC1 and PC2
cor.test(combenvir$PC2,combenvir$PC1) #highly correlated therefore cannot be includedi the same model

#Reorder speceis 
combenvir$Species<-factor(combenvir$Species, levels=c("Oyster", "Mud crab", "White shrimp", "Blue crab", "Pinfish"))

#Plot PC axes to contribution of benthic microalgae
p2<-ggplot(combenvir, aes(x=PC2, y=`Benthic algae`))+
  labs(y="Contribution of \n benthic microalgae production")+
   geom_point(shape=21, size=3, position=position_jitter(0.08), aes(fill=Site))+
    scale_fill_manual(values=c("#e66101", "#fdb863","#B2abd2","#5e3c99"))+
  facet_grid(~Species)+
  stat_smooth(method="lm", color="gray")+
  theme(legend.position="none")+
   theme(axis.text.x = element_text(angle = 45, hjust=1))

p1<-ggplot(combenvir, aes(x=PC1, y=`Benthic algae`))+
  labs(y="Contribution of \n benthic microalgae production")+
   geom_point(shape=21, size=3, position=position_jitter(0.08), aes(fill=Site))+
    scale_fill_manual(values=c("#e66101", "#fdb863","#B2abd2","#5e3c99"))+
  facet_grid(~Species)+
  stat_smooth(method="lm", color="gray")+
  theme(legend.position="none")+
   theme(axis.text.x = element_text(angle = 45, hjust=1))

#Combine PC plots into 1 figure
plot_grid(p1,p2, ncol=1, labels=c("A","B"),  label_size = 20)

#Open lme4 to run mixed models
library(lme4)

#Run linear mixed model with PC1 as fixed and species as a random variable
m1<-lmer(`Benthic algae`~PC1+(1|Species), data=combenvir, na.action=na.fail)
summary(m1)
MuMIn::r.squaredGLMM(m1)
MuMIn::dredge(m1)

#Run linear mixed model with PC2 as fixed and species as a random variable
m2<-lmer(`Benthic algae`~PC2+(1|Species), data=combenvir, na.action=na.fail)
MuMIn::r.squaredGLMM(m2)
MuMIn::dredge(m2)
#PC2 stronger model than using PC1 

#Visually examine output from PC2 model
library(visreg)

visreg(m2, type="contrast")

visreg(m2, type="conditional")

```


```{r}
library(hypervolume)
#Function to make random points of n length
# Generate random values from means and standard deviations.
# When end_points = TRUE, values are constrained to the interval defined by low and high.
# ID_rows and names can be supplied as either column names or column numbers, but they must match the same convention.
# df = data frame or tibble in which each row is a unique observation.
# ID_rows = vector of columns containing identifier information.
# names = column name or number for the measured variable name.
# mean = column name or number containing the mean.
# sd = column name or number containing the standard deviation.
# n = number of points to generate.
# z_score = TRUE/FALSE; if TRUE, z-score the values.
# end_points = TRUE/FALSE; if TRUE, generate between a high and low end point (for example, a 5% to 95% interval).
# low and high are required when end_points = TRUE.
# low = column name or number with the lower bound.
# high = column name or number with the upper bound.
HVvalues = function(df, ID_rows, names, mean, sd, n, z_score = T,
                    end_points = F, low = NULL, high = NULL){
  require(tidyverse)
  require(truncnorm)
  
  # check to see if information is needed to restrict where points are
  if (end_points){
    if (is_empty(df[,low]) | is_empty(df[,high])){
      return(cat('Warning: low and/or high columns not specified \n
                  Specific and run again or end_points = F \n'))
    }
  }
  
  # check to see if there are more 
  if (TRUE %in% duplicated(df[, c(ID_rows, names)])) {
    return(cat('Warning: some of the rows contain duplicated information \n
                make sure data is correct \n'))
  }
  
  # rename variables to make code work
  if (is.numeric(mean)) {
    names(df)[mean] = 'Mean'
  } else {
    df = df |>  rename(Mean = mean)
  }
  
  if (is.numeric(sd)) {
    names(df)[sd] = 'SD'
  } else {
    df = df |>  rename(SD = sd)
  }
  
  if (end_points) {
    if (is.numeric(low)) {
      names(df)[low] = 'lower'
    } else {
      df = df |>  rename(lower = low)
    }
    
    if (is.numeric(high)){
      names(df)[high] = 'upper'
    } else {
      df = df |>  rename(upper = high)
    }
  }
  
  # make sure the names is not numeric 
  if (is.numeric(names)) {
    names = names(df)[names]
  }
  
  labs = unique(as.vector(df[,names])[[1]])
  # generate random points within bounds
  if (end_points) {
    
    df_tot = df |> slice(rep(1:n(), each=n))|> 
      mutate(point = 
               truncnorm::rtruncnorm(1, a = lower, b = upper,
                                     mean = Mean, sd = SD)) |> 
      ungroup() |> 
      mutate(num = rep(1:n, times=nrow(df))) |>
      select(-Mean, -SD, -lower, -upper)|>
      pivot_wider(names_from = names, values_from = point)|> 
      select(-num)
  } else {
    # generate random points outside of bounds
    df_tot = df |> slice(rep(1:n(), each=n))|>
      mutate(point = 
               truncnorm::rtruncnorm(1, mean = Mean, sd = SD)) |> 
      ungroup() |> 
      mutate(num = rep(1:n, times=nrow(df))) |>
      select(-Mean, -SD)|>
      pivot_wider(names_from = names, values_from = point)|> 
      select(-num)
  }
  if (z_score) {
    df_tot = df_tot  |>  
      mutate(across(all_of(labs), scale))
  }
  
  return(df_tot)
  
}


# Load helper code for hypervolume analysis.
g = read_csv('Data/SIA/comb100.csv') |> 
  group_by(Site, Species, Source) |> 
  summarize(m = mean(Contr),
            sd = sd(Contr))


# Run hypervolume models for each site/species combination.
gb = HVvalues(df = g, ID_rows = c('Site','Species'),
              names = c('Source'), 
              mean = 'm', sd = 'sd', n = 100,
              end_points = F, 
              z_score = T) |> 
  group_by(Site, Species) |> 
  nest() |> 
  mutate(hv = map(data, ~hypervolume_gaussian(.x, name = paste(Site,Species,sep = '_'), # change to label columns in data
                                              samples.per.point = 1000,
                                              kde.bandwidth = estimate_bandwidth(.x), 
                                              sd.count = 3, 
                                              quantile.requested = 0.95, 
                                              quantile.requested.type = "probability", 
                                              chunk.size = 1000, 
                                              verbose = F)),
         hv_size = get_volume(hv[[1]]))


```

### Extract each hypervolume for each individual site and species

```{r}
#Oyster", "Mud crab", "White shrimp", "Blue crab", "Pinfish"
nco<-gb[[4]][[1]] #NC Oyster
get_volume(nco)
ncmc<-gb[[4]][[2]]#NC MC
get_volume(ncmc)
ncbc<-gb[[4]][[3]] #NC BC
get_volume(ncbc)
ncp<-gb[[4]][[4]]  #NC Pinfish
get_volume(ncp)
nio<-gb[[4]][[5]] #NI Oyster
get_volume(nio)
nimc<-gb[[4]][[6]] #NI MC
get_volume(nimc)
niws<-gb[[4]][[7]] #NI White shrimp
get_volume(niws)
nibc<-gb[[4]][[8]] #NI BC
get_volume(nibc)
nip<-gb[[4]][[9]] #NI Pinfish
get_volume(nip)
sio<-gb[[4]][[10]] #SI O
get_volume(sio)
simc<-gb[[4]][[11]] #SI MC
get_volume(simc)
siws<-gb[[4]][[12]] #SI WS
get_volume(siws)
sibc<-gb[[4]][[13]] #SI BC
get_volume(sibc)
sip<-gb[[4]][[14]] #SI Pinfish
get_volume(sip)
go<-gb[[4]][[15]] #GTM Oyster
get_volume(go)
gmc<-gb[[4]][[16]] #GTM MC
get_volume(gmc)
gws<-gb[[4]][[17]] #GTM WS
get_volume(gws)
gbc<-gb[[4]][[18]] #GTM BC
get_volume(gbc)
gp<-gb[[4]][[19]] #GTM Pinfish
get_volume(gp)


#Examine overlap statistics
#Create a hypervolume set for a site comparision for each species 
#Then look at the overlap statistics between each grouping
hv_ws1 <- hypervolume_set(gws, siws, check.memory=FALSE)
hypervolume_overlap_statistics(hv_ws1)

hv_ws2 <- hypervolume_set(gws, niws, check.memory=FALSE)
hypervolume_overlap_statistics(hv_ws2)

hv_ws3 <- hypervolume_set(siws, niws, check.memory=FALSE)
hypervolume_overlap_statistics(hv_ws3)


hv_o1 <- hypervolume_set(go, nco, check.memory=FALSE)
hypervolume_overlap_statistics(hv_o1)

hv_o2 <- hypervolume_set(sio, nio, check.memory=FALSE)
hypervolume_overlap_statistics(hv_o2)

hv_o3 <- hypervolume_set(sio, nco, check.memory=FALSE)
hypervolume_overlap_statistics(hv_o3)

hv_o4 <- hypervolume_set(nio, nco, check.memory=FALSE)
hypervolume_overlap_statistics(hv_o4)

hv_o5 <- hypervolume_set(nio, go, check.memory=FALSE)
hypervolume_overlap_statistics(hv_o5)

hv_o6 <- hypervolume_set(sio, go, check.memory=FALSE)
hypervolume_overlap_statistics(hv_o6)

hv_p1 <- hypervolume_set(gp, ncp, check.memory=FALSE)
hypervolume_overlap_statistics(hv_p1)

hv_p2 <- hypervolume_set(sip, nip, check.memory=FALSE)
hypervolume_overlap_statistics(hv_p2)

hv_p3 <- hypervolume_set(sip, ncp, check.memory=FALSE)
hypervolume_overlap_statistics(hv_p3)

hv_p4 <- hypervolume_set(nip, ncp, check.memory=FALSE)
hypervolume_overlap_statistics(hv_p4)

hv_p5 <- hypervolume_set(nip, gp, check.memory=FALSE)
hypervolume_overlap_statistics(hv_p5)

hv_p6 <- hypervolume_set(sip, gp, check.memory=FALSE)
hypervolume_overlap_statistics(hv_p6)

hv_mc1 <- hypervolume_set(gmc, ncmc, check.memory=FALSE)
hypervolume_overlap_statistics(hv_mc1)

hv_mc2 <- hypervolume_set(simc, nimc, check.memory=FALSE)
hypervolume_overlap_statistics(hv_mc2)

hv_mc3 <- hypervolume_set(simc, ncmc, check.memory=FALSE)
hypervolume_overlap_statistics(hv_mc3)

hv_mc4 <- hypervolume_set(nimc, ncmc, check.memory=FALSE)
hypervolume_overlap_statistics(hv_mc4)

hv_mc5 <- hypervolume_set(nimc, gmc, check.memory=FALSE)
hypervolume_overlap_statistics(hv_mc5)

hv_mc6 <- hypervolume_set(simc, gmc, check.memory=FALSE)
hypervolume_overlap_statistics(hv_mc6)

hv_bc1 <- hypervolume_set(gbc, ncbc, check.memory=FALSE)
hypervolume_overlap_statistics(hv_bc1)

hv_bc2 <- hypervolume_set(sibc, nibc, check.memory=FALSE)
hypervolume_overlap_statistics(hv_bc2)

hv_bc3 <- hypervolume_set(sibc, ncbc, check.memory=FALSE)
hypervolume_overlap_statistics(hv_bc3)

hv_bc4 <- hypervolume_set(nibc, ncbc, check.memory=FALSE)
hypervolume_overlap_statistics(hv_bc4)

hv_bc5 <- hypervolume_set(nibc, gbc, check.memory=FALSE)
hypervolume_overlap_statistics(hv_bc5)

hv_bc6 <- hypervolume_set(sibc, gbc, check.memory=FALSE)
hypervolume_overlap_statistics(hv_bc6)


set.seed(22)     
```
 
## Hypervolume visualization
```{r}
#Combine hypervolumes for each species (across all sites)
bc<-hypervolume_join(ncbc, nibc, sibc, gbc) #Blue crabs
oyst<-hypervolume_join(nco, nio, sio, go) #Oysters
mc<-hypervolume_join(ncmc, nimc, simc, gmc) #Mud crabs
pin<-hypervolume_join(ncp, nip, sip, gp) #Pinfish
ws<-hypervolume_join(niws, siws, gws) #White shrimp

#Blue crab
plot(bc, pairplot = T, colors=c("#5e3c99","#e66101", "#fdb863","#B2abd2" ), show.3d=FALSE,plot.3d.axes.id=NULL,
     show.axes=TRUE, show.frame=TRUE,
     show.random=T, show.density=TRUE,show.data=F,
     show.legend=T, 
     show.contour=F, contour.lwd= 2, 
     contour.type='alphahull', 
     contour.alphahull.alpha=0.25,
     contour.ball.radius.factor=1, 
     contour.kde.level=0.01,
     contour.raster.resolution=100,
     show.centroid=TRUE, cex.centroid=2,
     point.alpha.min=0.2, point.dark.factor=0.5,
     cex.random=0.5,cex.data=1,cex.axis=1.5,cex.names=2,cex.legend=2,
     num.points.max.data = 100000, num.points.max.random = 200000, reshuffle=TRUE,
     plot.function.additional=NULL,
     verbose=FALSE
)

#Mud crab
plot(mc, pairplot = T, colors=c("#5e3c99","#e66101", "#fdb863","#B2abd2" ), show.3d=FALSE,plot.3d.axes.id=NULL,
     show.axes=TRUE, show.frame=TRUE,
     show.random=T, show.density=TRUE,show.data=F,
     show.legend=T, 
     show.contour=F, contour.lwd= 2, 
     contour.type='alphahull', 
     contour.alphahull.alpha=0.25,
     contour.ball.radius.factor=1, 
     contour.kde.level=0.01,
     contour.raster.resolution=100,
     show.centroid=TRUE, cex.centroid=2,
     point.alpha.min=0.2, point.dark.factor=0.5,
     cex.random=0.5,cex.data=1,cex.axis=1.5,cex.names=2,cex.legend=2,
     num.points.max.data = 100000, num.points.max.random = 200000, reshuffle=TRUE,
     plot.function.additional=NULL,
     verbose=FALSE
)

#Oyster
plot(oyst, pairplot = T, colors=c("#5e3c99","#e66101", "#fdb863","#B2abd2" ), show.3d=FALSE,plot.3d.axes.id=NULL,
     show.axes=TRUE, show.frame=TRUE,
     show.random=T, show.density=TRUE,show.data=F,
     show.legend=T, 
     show.contour=F, contour.lwd= 2, 
     contour.type='alphahull', 
     contour.alphahull.alpha=0.25,
     contour.ball.radius.factor=1, 
     contour.kde.level=0.01,
     contour.raster.resolution=100,
     show.centroid=TRUE, cex.centroid=2,
     point.alpha.min=0.2, point.dark.factor=0.5,
     cex.random=0.5,cex.data=1,cex.axis=1.5,cex.names=2,cex.legend=2,
     num.points.max.data = 100000, num.points.max.random = 200000, reshuffle=TRUE,
     plot.function.additional=NULL,
     verbose=FALSE
)

#Pinfish
plot(pin, pairplot = T, colors=c("#5e3c99","#e66101", "#fdb863","#B2abd2" ), show.3d=FALSE,plot.3d.axes.id=NULL,
     show.axes=TRUE, show.frame=TRUE,
     show.random=T, show.density=TRUE,show.data=F,
     show.legend=T, 
     show.contour=F, contour.lwd= 2, 
     contour.type='alphahull', 
     contour.alphahull.alpha=0.25,
     contour.ball.radius.factor=1, 
     contour.kde.level=0.01,
     contour.raster.resolution=100,
     show.centroid=TRUE, cex.centroid=2,
     point.alpha.min=0.2, point.dark.factor=0.5,
     cex.random=0.5,cex.data=1,cex.axis=1.5,cex.names=2,cex.legend=2,
     num.points.max.data = 100000, num.points.max.random = 200000, reshuffle=TRUE,
     plot.function.additional=NULL,
     verbose=FALSE
)

#White shrimp
plot(ws, pairplot = T, colors=c("#e66101", "#fdb863","#B2abd2" ), show.3d=FALSE,plot.3d.axes.id=NULL,
     show.axes=TRUE, show.frame=TRUE,
     show.random=T, show.density=TRUE,show.data=F,
     show.legend=T, 
     show.contour=F, contour.lwd= 2, 
     contour.type='alphahull', 
     contour.alphahull.alpha=0.25,
     contour.ball.radius.factor=1, 
     contour.kde.level=0.01,
     contour.raster.resolution=100,
     show.centroid=TRUE, cex.centroid=2,
     point.alpha.min=0.2, point.dark.factor=0.5,
     cex.random=0.5,cex.data=1,cex.axis=1.5,cex.names=2,cex.legend=2,
     num.points.max.data = 100000, num.points.max.random = 200000, reshuffle=TRUE,
     plot.function.additional=NULL,
     verbose=FALSE
)

```

## Trophic level calculations

```{r}
#Using bootstrapped source contributions created above to calculate trophic levels of each species by site
Tlc<-comb100X%>% 
  #mutate(Site = fct_recode(Site, "NIWB" = "NI"))%>%
  group_by(Site, Species, Source)%>%
  mutate(rep=c(1:100))%>%
  spread(Source, Contr)
 
#Select source means N15 values 
sourcemeans<-read.csv("Data/SIA/MCsources.csv")%>%
  mutate(Site = fct_recode(Site, "NIWB" = "NI-WB"))%>%
  group_by(Site)%>%
  select(Source, Meand15N)%>%
  spread(Source, Meand15N)

sourcemeans<-sourcemeans%>%
  rename(Bmean=`Benthic algae`, Pmean=POM, Smean=Spartina)

#Combine bootstrapped file with source means
TLc1<-left_join(Tlc, sourcemeans, by="Site")


indv<-read.csv("Data/SIA/AllConsumers.csv")%>%
   mutate(Site = fct_recode(Site, "NIWB" = "NI-WB"))%>%
  group_by(Site, Species)%>%
  summarise(meand15N = mean(d15N))

TL1<-left_join(TLc1, indv, by=c("Site", "Species"))

#Calcualte TL using source contribtuions, source N15 means, and discrimination factor values
TL1<-TL1%>%
  mutate(TL = ((meand15N-((Bmean*`Benthic algae`)+(Pmean*POM)+(Smean*Spartina)))/3.3)+1)

```

## Trophic-level plots and statistical tests

```{r}
#Rename and reorder sites and species 
TL1<-TL1%>%mutate(Site = fct_recode(Site, "NI-WB"="NIWB"))
TL1$Site<-factor(TL1$Site, levels=c("NC", "NI-WB", "SI", "GTM"))
TL1$Species<-factor(TL1$Species, levels=c("Oyster", "Mud crab", "White shrimp", "Blue crab", "Pinfish"))

#Plot trophic levels 
ggplot(TL1, aes(x=Species, y=TL))+
  geom_jitter(color="gray80", size = 1)+
  geom_boxplot(aes(fill=Site))+
  scale_fill_manual(values=c("#e66101", "#fdb863","#B2abd2", "#5e3c99" ))+
  labs(y="Trophic level")+
  facet_wrap(~Site, ncol=2)+
  theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position="none")

#Calculate mean & SD for TL for tables
meanTLs<-TL1 %>%
  group_by(Site, Species)%>%
  summarize(meanTL=mean(TL), sdTL=sd(TL))


#Run Kruskal-Wallis and Dunn Posthoc tests
kruskal.test(TL~Site, data=TL1)
rstatix::dunn_test(data=TL1, TL~Site, p.adjust.method ="bonferroni")

#Use different package for Dunn Test for comparison
library(FSA)
dunnTest(TL~Site, data = TL1, method = "bonferroni")
```
